Co-citations in context

### 2019-06-03 cocit_r.tex
Update cocit_r.tex
Version prior to reorganization of Results section

Despite different methods used to generate expected values, both \citep{uzzi_atypical_2013} and \citep{boyack_vs_uzzi_2014} measured observed co-citation frequencies across the scientific literature without disciplinary constraints and subsequently used normalized frequencies to examine disciplinary subsets. In extending this prior work, we hypothesized that analysis of datasets focused in disciplinary areas would reduce model mis-specification since the pool of citations would be relevant to those disciplinary areas rather than being drawn from all disciplines. Accordingly, we used keyword searches of the scientific literature to create three citation networks themed around academic disciplines. Within these disciplinary frameworks, we calculated observed and expected co-citation frequencies using a refined random graph model and an efficient Monte Carlo simulation algorithm. The null model generated by citation shuffling in our study preserves the number of publications, the number of references group by year within each publication and, importantly, the disciplinary proportions of cited references in each dataset.

@@

\section{Materials and Methods}
\emph{Bibliographic data} We have previously developed ERNIE, an open source knowledge platform into which we parse the Web of Science (WoS) Core Collection~\citep{Keserci371955}. WoS data stored in ERNIE spans the period 1900-2019 and consists of over 72 million publications. For this study, we generated an analytical dataset from years 1985 to 2005 from ERNIE. The total number of publications in this dataset was just over 25 million publications (25,134,073).  For each of these years, we further restricted analysis to publications of type Article. Since WoS data also contains incomplete references or references that point at other indexes, we also considered only those references for which there were complete records~(\ref{tab:summary_data}). For example, WoS data for year 2005 contained 1,753,174 publications, which after restricting to type Article and considering only those references described above resulted in 916,573 publications, 6,095,594 unique references (set of references), and 17,167,347 total references (multiset of references). Given consistent trends in the data, we analyzed the two boundary years (1985 and 2005) and the mid-point (1995) performing 1,000 simulations for each dataset. \\
\emph{Disciplinary datasets} We constructed three disciplinary datasets based on keyword searches. (i) immunology (ii) metabolism (iii) applied physics. For the first two, rooted in biomedical research, we searched Pubmed for the term `immunology' or `metabolism' in the years 1985, 1995, and 2005~(\ref{tab:disc-datasets}). Pubmed IDs (pmids) returned were matched to WoS IDs (wos\_ids) and used to retrieve relevant articles. For the applied physics dataset, we directly searched subject labels in WoS for `applied physics'. We also examined publications in the five major research areas in the Web of Science; life sciences \& biomedicine, physical sciences, technology, social sciences, arts \& humanities using the extended subcategory classification of 153 sub-groups to categorize disciplinary composition of cited references in the datasets we studied.

@@

We investigated multiple definitions of hit articles, as did Uzzi et al., with hits defined as the 1\%, 2\%, 5\%, and 10\% top-cited articles.  Also, mention different thresholds, 1st and 10th, for novelty threshold.

@@

\section{Results}
\subsection{Chacko} In our study, we also use a Monte Carlo approach to simulate under a random graph model. A  principal consideration, however, was to restrict model misspecification arising from disciplinarily irrelevant references. We addressed this consideration by analyzing disciplinary subsets of the scientific literature, thereby restricting random selection of references to only those references in the disciplinary network being studied. We predicted that model misspecification as measured by the Kullback-Leibler (K-L) Divergence~\citep{kullback_information_1951} between observed and simulated frequencies in a disciplinary network would be less than the divergence for observed and simulated frequencies in the WoS superset. The results indicate that for the set of journals common to both a disciplinary network and the WoS superset~\ref{tab:KLD}, simulations under our model consistently have a lower K-L divergence compared to simulations that draw from the WoS superset (and its attendant substitutions that are ectopic with respect to field and discipline).

\begin{figure}
\includegraphics[width=0.5\hsize]{fold_effect.pdf}
\caption{Intra-network citation shuffling preserves disciplinary composition. Publications of type Article belonging to the WoS research areas (i) Arts and Humanities (ii) Life Sciences and Biomedicine were subject to a single shuffle of all their cited references using either the cited references in these networks as a source of random substitutions (bg\_local) or references from all articles in WoS (bg\_allWoS). The disciplinary composition of cited reference before and after shuffling was measured for each of 151 sub-groups (from the extended subject classification in WoS) in these two research areas, expressed as fold difference between citation counts grouped by subject for original (o) and shuffled (s) references using the formula (fold\_difference = $ifelse(o > s, o/s, s/o)$) and rounded to the nearest integer. A fold difference of 1 indicates that citation shuffling did not alter disciplinary composition. Data are shown for articles published in 1985. All four boxplots are generated from 151 observations. Min \& Max values for fold differences were as follows. (i) Arts \& Humanities. bg\_local (1,2), bg\_allWoS (1, 488) (ii) Life Sciences \& Biomedicine. bg\_local(1,1), bg\_allWos (1, 352) Note y-axis: $log_2$ scale.} 
\end{figure}

@@

>>>>

Despite different methods used to generate expected values, both \citep{uzzi_atypical_2013} and \citep{boyack_vs_uzzi_2014} measured observed co-citation frequencies across the scientific literature without disciplinary constraints and subsequently used normalized frequencies to examine disciplinary subsets. In extending this prior work, we hypothesized that analysis of datasets focused in disciplinary areas would reduce model mis-specification since the pool of citations would be relevant to those disciplinary areas rather than being drawn from all disciplines. Accordingly, we used keyword searches of the scientific literature to create three citation networks themed around academic disciplines. Within these disciplinary frameworks, we calculated observed and expected co-citation frequencies using a refined random graph model and an efficient Monte Carlo simulation algorithm. The null model generated by citation shuffling in our study preserves the number of publications, the number of references group by year within each publication and, importantly, the disciplinary proportions of cited references in each dataset. Our analysis (see Results section) demonstrates the model misspecification prior studies and its resolution in our approach.

@@

\section{Materials and Methods}
\emph{Bibliographic data} We have previously developed ERNIE, an open source knowledge platform into which we parse the Web of Science (WoS) Core Collection~\citep{Keserci371955}. WoS data stored in ERNIE spans the period 1900-2019 and consists of over 72 million publications. For this study, we generated an analytical dataset from years 1985 to 2005 from ERNIE. The total number of publications in this dataset was just over 25 million publications (25,134,073).  For each of these years, we further restricted analysis to publications of type Article. Since WoS data also contains incomplete references or references that point at other indexes, we also considered only those references for which there were complete records~(\ref{tab:summary_data}). For example, WoS data for year 2005 contained 1,753,174 publications, which after restricting to type Article and considering only those references described above resulted in 916,573 publications, 6,095,594 unique references (set of references), and 17,167,347 total references (multiset of references). Given consistent trends in the data, we analyzed the two boundary years (1985 and 2005) and the mid-point (1995) performing 1,000 simulations for each dataset.

\emph{Disciplinary datasets} We constructed three disciplinary datasets based on keyword searches. (i) immunology (ii) metabolism (iii) applied physics. For the first two, rooted in biomedical research, we searched Pubmed for the term `immunology' or `metabolism' in the years 1985, 1995, and 2005~(\ref{tab:disc-datasets}). Pubmed IDs (pmids) returned were matched to WoS IDs (wos\_ids) and used to retrieve relevant articles. For the applied physics dataset, we directly searched subject labels in WoS for `applied physics'. We also examined publications in the five major research areas in the Web of Science; life sciences \& biomedicine, physical sciences, technology, social sciences, arts \& humanities using the extended subcategory classification of 153 sub-groups to categorize disciplinary composition of cited references in the datasets we studied.

@@

\emph{This comment needs to be worked into the M\&M section in an appropriate place, with any other specifics about the experiment parameters} We investigated multiple definitions of hit articles, as did Uzzi et al., with hits defined as the 1\%, 2\%, 5\%, and 10\% top-cited articles.  Also, mention different thresholds, 1st and 10th, for novelty threshold.

@@

\section{Results}

\subsection{Model Misspecification and the Attributes of Disciplinary Context}

\subsection{Novelty and Conventionality as Determinants of Impact in Disciplinary Contexts}

\subsection{Chacko} In our study, we also use a Monte Carlo approach to simulate under a random graph model. A  principal consideration, however, was to restrict model misspecification arising from disciplinarily irrelevant references. We addressed this consideration by analyzing disciplinary subsets of the scientific literature, thereby restricting random selection of references to only those references in the disciplinary network being studied. We predicted that model misspecification as measured by the Kullback-Leibler (K-L) Divergence~\citep{kullback_information_1951} between observed and simulated frequencies in a disciplinary network would be less than the divergence for observed and simulated frequencies in the WoS superset. The results indicate that for the set of journals common to both a disciplinary network and the WoS superset~\ref{tab:KLD}, simulations under our model consistently have a lower K-L divergence compared to simulations that draw from the WoS superset (and its attendant substitutions that are ectopic with respect to field and discipline).

\begin{figure}
\centering
\includegraphics[width=0.5\hsize]{fold_effect.pdf}
\caption{Intra-network citation shuffling preserves disciplinary composition. Publications of type Article belonging to the WoS research areas (i) Arts and Humanities (ii) Life Sciences and Biomedicine were subject to a single shuffle of all their cited references using either the cited references in these networks as a source of random substitutions (bg\_local) or references from all articles in WoS (bg\_allWoS). The disciplinary composition of cited reference before and after shuffling was measured for each of 151 sub-groups (from the extended subject classification in WoS) in these two research areas, expressed as fold difference between citation counts grouped by subject for original (o) and shuffled (s) references using the formula (fold\_difference = $ifelse(o > s, o/s, s/o)$) and rounded to the nearest integer. A fold difference of 1 indicates that citation shuffling did not alter disciplinary composition. Data are shown for articles published in 1985. All four boxplots are generated from 151 observations. Min \& Max values for fold differences were as follows. (i) Arts \& Humanities. bg\_local (1,2), bg\_allWoS (1, 488) (ii) Life Sciences \& Biomedicine. bg\_local(1,1), bg\_allWos (1, 352) Note y-axis: $log_2$ scale.} 
\end{figure}

@@

George's point from Slack conversation: ``A concern that surfaced on the drive are the definitions of conventionality and novelty. On the one hand we use them- on the other hand you've also pointed out that it isn't that simple.''
